### 2020-05-30 writeup/setup.tex
sec 2

We describe the bias-variance tradeoff of our setting as a warmup.
Recall that we have $t$ labeled tasks available, denoted by $(X_1, Y_1), (X_2, Y_2), \dots, (X_t, Y_t)$, where $X_i\in\real^{n_i\times p}$ and $Y_i\in\real^{n_i}$ for $1\le i\le t$.
%Following \cite{HMRT19,BLLT20}, we assume that for each task $i = 1,2,\dots,t$,  every feature vector is generated as $x = \Sigma_i^{1/2} z$, where $z\in\real^p$ is a random vector with i.i.d. entries of mean zero and unit variance and $\Sigma_i\in\real^{p\times p}$ is a positive semidefinite matrix.
Without loss of generality, let the $t$-th task denote the target task.
For an estimator $\hat{\beta}\in\real^p$, the test error of the target task is defined as
	\begin{align*}
		\te_t(\hat{\beta}) \define \exarg{z}{\exarg{\varepsilon_t}{({(\Sigma_t^{1/2} z)}^{\top}\hat{\beta} - {(\Sigma_t^{1/2})}^{\top}\beta_t)^2}}
		= \exarg{\varepsilon_t}{(\hat{\beta} - \beta_t)^{\top}\Sigma_t(\hat{\beta} - \beta_t)}.
	\end{align*}
The single-task estimator $\hat{\beta}_t^{\STL}$ is given by $(X_t^{\top}X_t)^{-1}X_t^{\top}Y_t$.
The bias-variance trade-off \cite{HTF09} says
	\[ \te_t(\hat{\beta}) =
		\bignorm{\exarg{\varepsilon_t}{\hat{\beta}} - \beta_t}^2 + \exarg{\varepsilon_t}{\bignorm{\hat{\beta} - \exarg{\varepsilon_t}{\hat{\beta}}}^2}. \]
In order to study the trade-off between model-shift bias and variance reduction, we need tight concentration bounds to quantify both effects.
For this purpose, we consider the high-dimensional regime where $n_i$ is a fixed constant $\rho_i > 1$ times $p$ for every $1\le i\le t$, and $p$ is large.
%Recall that $n_i = \rho_i \cdot p$ and we assume $\rho_i > 1$ is a fixed constant for every $1\le i\le t$.

@@

>>>>

We describe the bias-variance tradeoff of our setting more formally.
Recall that we have $t$ labeled tasks available, denoted by $(X_1, Y_1), (X_2, Y_2), \dots, (X_t, Y_t)$, where $X_i\in\real^{n_i\times p}$ and $Y_i\in\real^{n_i}$ for $1\le i\le t$.
%Following \cite{HMRT19,BLLT20}, we assume that for each task $i = 1,2,\dots,t$,  every feature vector is generated as $x = \Sigma_i^{1/2} z$, where $z\in\real^p$ is a random vector with i.i.d. entries of mean zero and unit variance and $\Sigma_i\in\real^{p\times p}$ is a positive semidefinite matrix.
Without loss of generality, let the $t$-th task denote the target task.
For an estimator $\hat{\beta}\in\real^p$, we define the out-of-sample (prediction) loss as
	\begin{align*}
		\te_t(\hat{\beta}) \define \exarg{z}{\exarg{\varepsilon_t}{({(\Sigma_t^{1/2} z)}^{\top}\hat{\beta} - {(\Sigma_t^{1/2})}^{\top}\beta_t)^2}}
		= \exarg{\varepsilon_t}{(\hat{\beta} - \beta_t)^{\top}\Sigma_t(\hat{\beta} - \beta_t)}.
	\end{align*}
The single-task estimator $\hat{\beta}_t^{\STL}$ is given by $(X_t^{\top}X_t)^{-1}X_t^{\top}Y_t$.
The bias-variance trade-off \cite{HTF09} says
	\[ \te_t(\hat{\beta}) =
		\bignorm{\exarg{\varepsilon_t}{\hat{\beta}} - \beta_t}^2 + \exarg{\varepsilon_t}{\bignorm{\hat{\beta} - \exarg{\varepsilon_t}{\hat{\beta}}}^2}. \]
In order to study the trade-off between model-shift bias and variance reduction, we need tight concentration bounds to quantify both effects.
For this purpose, we consider the high-dimensional regime where $n_i$ is a fixed constant $\rho_i > 1$ times $p$ for every $1\le i\le t$, and $p$ is large.
%Recall that $n_i = \rho_i \cdot p$ and we assume $\rho_i > 1$ is a fixed constant for every $1\le i\le t$.

@@

\section{A Technical Tool to Quantify the Bias-Variance Trade-off}
\label{sec_main}

%We revisit the classical bias-variance trade-off in the context of multi-task learning for linear regressions, in particular the setup of equation \eqref{eq_mtl}.
We develop a technical tool to derive $\te(\hat{\beta}_t^{\MTL})$ that only depends on the qualities of task data such as data sizes and covariance matrices, for two tasks with general covariances.
Then, we extend the result to more than two tasks that share the same features but have different labels.
Finally, we show a sharp bias-variance tradeoff for transfer learning settings.
%For the high-dimensional setting where the data size of the target task is a small constant times dimension $p$, we quantify the trade-off as a function of the task models $\set{\beta_i}_{i=1}^t$, data size $\set{n_i}_{i=1}^t$, and covariance matrices $\set{\Sigma_i}_{i=1}^t$.

\textbf{Multi-task learning.} applies to the setting of two tasks where their covariance matrices may be arbitrarily different.
%The technical crux of our approach is to derive the asymptotic limit of $\te(\hat{\beta}_t^{\MTL})$ in the high-dimensional setting, when $p$ approaches infinity.
We derive a precise limit of $\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}$, which is a deterministic function that only depends on $\Sigma_1, \Sigma_2$ and $n_1/p, n_2/p$ (see Lemma \ref{lem_cov_shift} in Appendix \ref{app_proof_main} for the result).
Based on the result, we show how to determine positive versus negative transfer as follows.

\begin{theorem}[Informal]\label{thm_main_informal}
	Let $X_i \in\real^{n_i \times p}$ and $Y_i = X_i\beta_i + \varepsilon_i$, for $i = 1, 2$.
	Suppose that $n_1 = \rho_1 p$ and $n_2 = \rho_2 p$, where $\rho_1>1$ and $\rho_2 >1$ are fixed constants.
	There exists two deterministic functions $\Delta_{\beta}$ and $\Delta_{\vari}$ and a small deterministic error $\delta$ that only depend on $\set{\hat{v}, \Sigma_1^{}, n_1, n_2, \beta_1, \beta_2}$ such that
	\begin{itemize}
		\item If $\Delta_{\vari} - \Delta_{\beta} \ge \delta$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\Delta_{\vari} - \Delta_{\beta} \le \delta$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\end{itemize}
\end{theorem}

Theorem \ref{thm_main_informal} shows nearly tight bounds on the trade-off between model-shift bias and variance reduction.
%determined by the covariate shift matrix and the model shift.
The bounds get tighter and tighter as $\rho_1$ increases.
While the general form of $\Delta_{\vari}$ and $\Delta_{\beta}$ can be quite complex, we will show that they provide nice interpretation for simplified settings later on in Section \ref{sec_insight}.

We describe an overview of the proof of Theorem \ref{thm_main_informal}.
By comparing the bias and variance of $\te_t(\hat{\beta}_t^{\MTL})$ to $\te_t(\hat{\beta}_t^{\STL})$, we observe the following two effects, respectively.
%We begin by observing that the test error of $\hat{\beta}_t^{\MTL}$ consists of two parts.
We term equation \eqref{eq_te_model_shift} as \textit{model-shift bias}, which captures how similar $\beta_1$ and $\beta_2$ are.
This part introduces a negative effect to multi-task learning.
The scaling term $\hat{v}$ corresponds to the intuition that the shared module $B$ learns a subspace for both tasks.
The output layer of each task scales the direction of $B$ suitably to fit the task.

Next, it is not hard to verify that equation \eqref{eq_te_var}, the variance of $\hat{\beta}_t^{\MTL}$, is always smaller than the variance of $\hat{\beta}_t^{\STL}$.
This part introduces a positive variance reduction effect to performing multi-task learning.
Hence, whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$ is determined by the trade-off between two effects:
(i) the positive effect from variance reduction;
(ii) the negative effect from model shift bias.
A formal version of Theorem \ref{thm_main_informal} is presented in Theorem \ref{thm_model_shift} and its proof is presented in Appendix \ref{app_proof_main}.

\textbf{Extension.} The above result also applies to the setting of more than two tasks with the same covariates.
%We extend the above result to any number of tasks that have the same covariates.
%In this section we consider the setting with $k$ many that have the same covariates.
Since the tasks all have the same number of datapoints and covariance matrix, the trade-off between model shift bias and variance will be captured by their task models $\set{\beta_i}_{i=1}^k$.
%For this setting, we derive solutions for the multi-task training and the transfer learning setting that match our insights qualitatively from Section \ref{sec_denoise}.
%Let $B^{\star} = [\beta_1, \beta_2, \dots, \beta_k] \in\real^{p\times k}$ denote the underlying task model parameters.
We derive a similar trade-off between model shift bias and variance reduction for this setting as well.
The formal statement is stated in Theorem \ref{thm_many_tasks} and its proof can be found in Appendix \ref{app_proof_many_tasks}.

\textbf{Transfer learning.}
We extend the intuition behind Theorem \ref{thm_main_informal} to transfer learning settings.
We provide an analysis of the transfer function of Taskonomy \cite{ZSSGM18} using our setup.
Specifically, the source task encoder consists of the representations learnt from one or more source tasks.
The transfer function then tries to fit the target task data to the source task encoder.
For more details, we refer the reader to Figure 4 in Taskonomy.

We map the procedure to our setup as follows.
First, we obtain the single-task estimator $\hat{\beta}_i$ from the source tasks, for $1\le i \le t-1$.
This forms the shared representation $B = [\hat{\beta}_1,\hat{\beta}_2,\dots,\hat{\beta}_{t-1}]$.
Then, we learn the output layer $W_t$ on the target task by minimizing the following objective
\begin{align}
	g(W_t) = \bignorm{X_t B W_t - Y_t}^2.
\end{align}
After solving $W_t$, we use $\hat{\beta}_t^{\TL} = B W_t$ as the estimator for the target task.
By comparing $\te(\hat{\beta}_t^{\TL})$ to $\te(\hat{\beta}_t^{\STL})$, we observe a similar trade-off between model-shift bias and variance reduction for this setting.
%We use our tools to compare $\te(B W_t)$ to $\te(\beta_t^{\STL})$.
The formal statement is presented in Theorem \ref{prop_taskonomy} and its proof in Appendix \ref{app_proof_sec4}.
